---
title: "Create PMML export based on logistic regression model leveraging Framingham heart study data"
author: "Alexander Lemm"
date: "27. March 2015"
output: html_document
---

```{r warning=FALSE, message=FALSE}

library(pmml)
library(caTools)
library(ROCR)

```



## Read and process data 

```{r}

framingham <- read.csv("./data/framingham.csv")

dim(framingham)
sum(complete.cases(framingham))

framingham <- framingham[complete.cases(framingham), ]

```

What is the accuracy of the baseline model?

```{r}

tab <- table(framingham$TenYearCHD)
tab

baseline <- tab[1]/sum(tab)
baseline

```

Split the data into training and test set:

```{r}

set.seed(37)

split <- sample.split(framingham$TenYearCHD, SplitRatio = 0.65)

fram_train <- framingham[which(split), ]
fram_test <- framingham[which(!split), ]

```



## Build logistic regression models and apply them on training set 

## First model

Build model including all available independent variables:

```{r}
fram_log_complete <- glm(TenYearCHD ~ ., family = binomial, data = fram_train)
summary(fram_log_complete)

```

The following variables are significant: `male`, `age`, `cigsPerDay`, `totChol`,
`sysBP` and `glucose`.

In addition, we observe an AIC of `r summary(fram_log_complete)$aic`.


Next, we evaluate the in-sample accuracy.

```{r}

predict_train1 <- predict(fram_log_complete, type = "response")
t1_train <- table(fram_train$TenYearCHD, predict_train1 >= 0.5)
t1_train

sum(diag(t1_train))/sum(t1_train)

```

### Second model

Next, we build a second model using some of the significant variables plus those identified as risk factors for corony heart diseases in other studies.

```{r}

fram_log2 <- glm(TenYearCHD ~ male + age + currentSmoker + totChol + sysBP +
                   glucose, family = 'binomial', data = fram_train)
summary(fram_log2)

```

All independent variables are significant  whereas 4 out of 6 are even highly statistically significant. All $\beta$ coefficients show a positive sign.

```{r}

predict_train2 <- predict(fram_log2, type = "response")

t2_train <- table(fram_train$TenYearCHD, predict_train2 >= 0.5)
t2_train
sum(diag(t2_train))/sum(t2_train)

```

We will only apply the second model on our test set even though the first one possesses a slightly higher in-sample accuracy and a slightly higher AIC.


## Apply final model on test set 

```{r}

predict_test <- predict(fram_log2, type = "response", newdata = fram_test)
summary(predict_test)
t1_test <- table(fram_test$TenYearCHD, predict_test >= 0.5)
t1_test
acc <- sum(diag(t1_test))/sum(t1_test)
acc

```

The result shows that our logistic regression model (`r acc`) has a slight edge over the baseline model (`r baseline`) in terms of overall accuracy. 

Let's have a look at the Receiver Operator Characteristic curve:

```{r}

rocr_pred <- prediction(predict_test, fram_test$TenYearCHD)
rocr_perf <- performance(rocr_pred, "tpr", "fpr")
plot(rocr_perf, colorize = TRUE, print.cutoffs.at = seq(0, 1, 0.1), 
     text.adj = c(-0.1, 1))


```

What about the AUC?

```{r}

rocr_auc <- as.numeric(performance(rocr_pred, "auc")@y.values)
rocr_auc

```


## Export PMML file

```{r results='hide'}

glm_pmml <- pmml(fram_log2, model.name = "CHD_prediction_model")
xmlFile <- file.path(getwd(),"files/chd_risk_glm.xml")
saveXML(glm_pmml,xmlFile)

```


